---
title: "Comparing Multiple Forecasters with SMCS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparing Multiple Forecasters with SMCS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4
)

```

```{r setup}
library(seqcomp)

```

## Overview

While `compare_forecasts()` evaluates exactly two models, real-world applications often require comparing an arbitrary number of candidate models ($m \ge 2$) simultaneously.

Evaluating multiple models sequentially introduces a multiple-testing problem: if you run enough pairwise comparisons over enough time steps, you will eventually find false "statistically significant" differences by pure chance.

To solve this, `seqcomp` implements **Sequential Model Confidence Sets (SMCS)** (Arnold et al., 2026). The SMCS is the set of models that have not yet been confidently beaten by any other model in the candidate pool. The package guarantees that the true best model(s) will remain in the SMCS with probability at least $1 - \alpha$ over the entire evaluation period. *(If you are only comparing two forecasters, we recommend starting with the [introductory vignette](seqcomp.html).)*

## Two Notions of Superiority

When comparing multiple models, `seqcomp` tracks two different definitions of what makes a model "superior":

1. **The Strong Null (`smcs_strong`)**: A model is strongly superior if it performs at least as well as every other model *at every single time step*. If a model is consistently good but has occasional severe failures, it will be eliminated from this set.
2. **The Weak Null (`smcs_weak`)**: A model is weakly superior if it performs at least as well as every other model *on average* over time. This allows models to remain in the set even if they occasionally issue bad forecasts, provided their long-term average remains dominant.

## A simple multi-model example

Let's generate a sequence of binary outcomes and three candidate forecasters:

* `M1`: A highly accurate forecaster.
* `M2`: A mediocre forecaster that guesses randomly.
* `M3`: A terrible forecaster that is consistently wrong.

```{r}
set.seed(2026)
T_sim <- 300
y <- rbinom(T_sim, size = 1, prob = 0.7)

# Create 3 forecasters
forecasts <- matrix(0, nrow = T_sim, ncol = 3)
colnames(forecasts) <- c("M1", "M2", "M3")

forecasts[, "M1"] <- ifelse(y == 1, 0.8, 0.2)   # Highly accurate
forecasts[, "M2"] <- runif(T_sim, 0.4, 0.6)     # Mediocre/random
forecasts[, "M3"] <- ifelse(y == 1, 0.2, 0.8)   # Consistently wrong

head(forecasts)

```

## Compare the forecasts

The easiest workflow for evaluating three or more models is to use `compare_multiple_forecasts()`.

```{r}
multi_cmp <- compare_multiple_forecasts(
  forecasts = forecasts,
  outcomes = y,
  scoring_rule = "brier",
  alpha = 0.05
)

# Printing the object provides a clean summary of which models 
# were excluded and when they first dropped out of the set.
multi_cmp

```

While the printed summary gives a quick overview, the underlying object is a list containing the full time-indexed matrices for further analysis(alongside metadata such as `alpha` and `scoring_rule`):

* `multi_cmp$scores`: The evaluated pointwise scores for each model.
* `multi_cmp$smcs_strong`: A logical matrix tracking whether each model belongs to the strong-null SMCS at time $t$.
* `multi_cmp$smcs_weak`: A logical matrix tracking whether each model belongs to the weak-null SMCS at time $t$.

## Interpreting the SMCS output

`smcs_strong()` maintains a **running intersection** over time: once a model
is shown to violate the null against some competitor, it is excluded
permanently. This matches the strong/uniformly-weak SMCS construction of
Arnold, Gavrilopoulos, Schulz and Ziegel (2026, Section 3.2).

The underlying hypothesis of `smcs_weak()` (the *time-varying* weak null, 
Section 3.3) is defined *at* each time $t$, not cumulatively, so a
model's average performance can legitimately recover after a bad stretch.
`smcs_weak()` therefore returns `smcs`, which can re-admit models that were 
previously excluded.

Let's look at the strong SMCS at the end of the evaluation:

```{r}
tail(multi_cmp$smcs_strong)

```

We can visualize how the confidence set shrinks over time by plotting whether a model is `TRUE` (Included) or `FALSE` (Excluded):

```{r}
par(mfrow = c(3, 1), mar = c(2, 4, 2, 1))
colors <- c("blue", "gray", "red")

for (i in 1:3) {
  plot(
    1:T_sim, multi_cmp$smcs_strong[, i], 
    type = "s", col = colors[i], lwd = 2,
    ylim = c(-0.1, 1.1), yaxt = "n", ylab = "In SMCS?",
    main = paste("Model:", colnames(forecasts)[i])
  )
  axis(2, at = c(0, 1), labels = c("Excluded", "Included"), las = 2)
}
par(mfrow = c(1, 1))

```

Notice how quickly the set shrinks:

1. `M3` (the terrible model) is confidently beaten by `M1` almost immediately and drops out of the set.
2. `M2` (the mediocre model) survives a bit longer, but as evidence accumulates, it too is permanently excluded.
3. `M1` (the true best model) remains in the SMCS for the entire duration.

## Using lower-level functions directly

Just like the two-model case, `seqcomp` exposes the underlying multi-model primitives if you need custom bounds, different scoring rules, or adaptive betting fractions.

You can compute your own score matrices and pass them directly to `smcs_strong()` or `smcs_weak()`.

```{r}
# Manually compute Brier scores
scores_mat <- matrix(0, nrow = T_sim, ncol = 3)
for(i in 1:3) scores_mat[, i] <- brier_score(forecasts[, i], y)

# Construct the Weak SMCS directly
# Brier score differences are bounded in [-1, 1], so we use c_param = 2
res_weak <- smcs_weak(
  scores = scores_mat,
  alpha = 0.05,
  cs_method = "bernstein",
  c_param = 2
)

tail(res_weak$smcs)

```

## Conditionally bounded scores (Tick loss)

Some scoring rules, like tick loss for quantile forecasting, are unbounded globally but bounded *conditionally* based on the distance between the forecasts.

For these rules, the strong SMCS requires a dynamically updating 3D array of bounds and betting fractions. The wrapper `compare_multiple_forecasts()` handles this automatically when `scoring_rule = "tick"`, or you can construct the arrays manually using `build_quantile_betting_arrays()`.

```{r}
set.seed(7)
tau <- 0.5

truth <- rnorm(T_sim)
q_forecasts <- cbind(
  M1 = truth + rnorm(T_sim, sd = 0.1),    # accurate
  M2 = rep(0, T_sim),                     # uninformative
  M3 = truth + rnorm(T_sim, sd = 0.1) + 1 # biased
)

tick_cmp <- compare_multiple_forecasts(
  forecasts    = q_forecasts,
  outcomes     = truth,
  scoring_rule = "tick",
  tau          = tau,
  alpha        = 0.05
)

tail(tick_cmp$smcs_strong)
```

`smcs_weak` is `NULL` here, since tick-loss score differences are only
conditionally bounded, not uniformly bounded, and `smcs_weak()` currently
requires a fixed uniform bound.
